---
title: "poleis"
output: html_document
---

Script to generate environmental inputs for the Greek poleis case study.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(stars)
library(tidygraph)
```


```{r}
poleis <- read_tsv('~/Downloads/868961/RW109figsharev2/RW109b_verticesbasic4.txt') %>%
  st_as_sf(coords = c('Longitude_E', 'Latitude_N'), crs = 4326) %>%
  st_transform(32634)
```

```{r}
radius <- 2500
hex <- st_bbox(poleis) %>%
    st_make_grid(cellsize = radius * sqrt(3), 
                 square = FALSE, 
                 flat_topped = FALSE) %>% # st_make_grid wants the short diagonal
    st_sf() %>%
  mutate(id = 1:n())

hex_geo <- st_transform(hex, 4326)
```

```{r}
site_count <- st_join(poleis, hex, join = st_within) %>%
  as_tibble() %>%
  group_by(id) %>%
  summarize(count = n()) %>%
  left_join(hex, ., by = 'id') %>%
  replace_na(list(count = 0))

plot(site_count)
```



```{r, message = FALSE}
npp <- read_stars(c('../urban-persistence/CHELSA_bio10_01.tif', 
                    '../urban-persistence/CHELSA_bio10_12.tif')) %>% 
  setNames(c('temperature', 'precipitation')) %>%
  mutate(temperature = temperature / 10, # temperature is in degrees C * 10
         ### the miami model
         npp_prec = 3 * (1 - exp(-0.000664 * precipitation)),
         npp_temp = 3 / (1 + exp(1.315 - 0.119 * temperature)),
         npp = pmin(npp_prec, npp_temp)) %>% # kilograms of dry matter / square meter / year
  select(temperature, precipitation, npp) %>%
  st_crop(st_bbox(hex_geo)) %>%
  aggregate(hex_geo, mean, na.rm = TRUE) %>% 
  st_transform(hex) %>%
  st_as_sf()
```


```{r, message = FALSE}
elev <- read_stars('~/Downloads/EarthEnv-DEM90_N35E020/EarthEnv-DEM90_N35E020.bil') %>%
  st_crop(hex_geo)

land_mask <- aggregate(elev >= 1, hex_geo, mean, na.rm = TRUE) %>% 
  st_as_sf() %>%
  st_transform(st_crs(hex)) %>%
  rename(land_mask = EarthEnv.DEM90_N35E020.bil)

arable <- elev %>%
  as('Raster') %>%
  raster::terrain(unit = 'degrees') %>%
  st_as_stars() %>%
  aggregate(hex_geo, FUN = function(x) sum(x <= 5, na.rm = TRUE) / length(x)) %>%
  st_transform(hex) %>% 
  st_as_sf() %>%
  st_join(land_mask, join = st_equals) %>%
  mutate(arable = as.numeric(slope * st_area(.) * land_mask),
         # mask out cells with less than half land (~8km2)
         arable = if_else(land_mask >= 0.5, arable, NA_real_))
```

multiply by to .25 to get aboveground biomass and grain. now edible grain in kg/m2, multiply times arable land to get total kg per cell, divide by 300 to get people it can support and * 0.50 for biennial fallow too. this is the free food from the agricultural population, so multiply by 0.1 as the "tax" that's actually available to move around
```{r, message = FALSE}
env <- st_join(arable, npp, join = st_equals) 

food <- transmute(env, carrying_capacity = npp * 0.25 * arable / 300 / 2 * 0.1)
```

```{r}
plot(env)
```

```{r}
plot(food)
```


```{r}
pts <- food %>%
  mutate(id = 1:n()) %>%
  filter(!is.na(carrying_capacity)) %>% 
  st_centroid()

plot(pts)
```

```{r}
settlements <- pts %>%
  st_distance() %>%
  units::set_units('km') %>%
  units::drop_units() %>% 
  replace(. == 0, 99999) %>%
  as_tbl_graph() %>%
  mutate(id = pts$id,
         population = 15,
         food = pts$carrying_capacity,
         eq = 0,
         x = st_coordinates(pts)[,1], 
         y = st_coordinates(pts)[,2]) %E>%
  rename(distance = weight) %>%
  # set self ties to half the cell radius, in km
  mutate(distance = if_else(distance == 99999, radius/2000, distance),
         trade_flow = 0,
         migrant_flow = 0)
```

```{r}
saveRDS(settlements, 'greece_settlements')
saveRDS(food, 'greece_env')
```

```{r}
polis_sim <- accumulate(1:1000,
                   ~run_sim(., beta1 = 3, beta2 = 6, alpha1 = 1.05, alpha2 = 0, nu = 0.05), 
                   .init = prune(settlements, 6)) 
```

```{r}
library(gghighlight)

site_names <- st_join(poleis, hex, join = st_within) %>%
  as_tibble() %>%
  select(id, Name)

polis_sim %>%
  map_dfr(as_tibble, 'nodes', .id = 'time') %>%
  mutate(time = as.numeric(time)) %>%
  left_join(site_names) %>%
  ggplot(aes(time, population)) +
  geom_line(aes(group = id)) +
    gghighlight(Name %in% c('Athens','Thebes', 'Argos', 'Korinth'),
                label_key = Name,
                unhighlighted_params = list(alpha = 0.3)) +
  theme_bw() +
  labs(x = 'Years', y = 'Population')
```

```{r}
rm(polis_sim);gc()
polis_sim_cc <- accumulate(1:1000,
                   ~run_sim(., beta1 = 3, beta2 = 6, alpha1 = 1.05, alpha2 = 0, nu = 0.05), 
                   .init = prune(settlements, 6) %N>% mutate(population = food)) 
```



```{r}


polis_sim[[800]] %N>%
  as_tibble %>%
  filter(population > 1) %>%
  pull(population)  %>% hist
```

