---
title: "poleis"
output: html_document
---

Script to generate environmental inputs for the Greek poleis case study.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(stars)
library(tidygraph)
library(ggrepel)

library(gghighlight)
```


```{r}
poleis <- read_tsv('~/Downloads/868961/RW109figsharev2/RW109b_verticesbasic4.txt') %>%
  st_as_sf(coords = c('Longitude_E', 'Latitude_N'), crs = 4326) %>%
  st_transform(32634)
```

```{r}
radius <- 2500
hex <- st_bbox(poleis) %>%
    st_make_grid(cellsize = radius * sqrt(3), 
                 square = FALSE, 
                 flat_topped = FALSE) %>% # st_make_grid wants the short diagonal
    st_sf() %>%
  mutate(id = 1:n())

hex_geo <- st_transform(hex, 4326)
```

```{r}
site_count <- st_join(poleis, hex, join = st_within) %>%
  as_tibble() %>%
  group_by(id) %>%
  summarize(count = n()) %>%
  left_join(hex, ., by = 'id') %>%
  replace_na(list(count = 0))

plot(site_count)
```



```{r, message = FALSE}
npp <- read_stars(c('../urban-persistence/CHELSA_bio10_01.tif', 
                    '../urban-persistence/CHELSA_bio10_12.tif')) %>% 
  setNames(c('temperature', 'precipitation')) %>%
  mutate(temperature = temperature / 10, # temperature is in degrees C * 10
         ### the miami model
         npp_prec = 3 * (1 - exp(-0.000664 * precipitation)),
         npp_temp = 3 / (1 + exp(1.315 - 0.119 * temperature)),
         npp = pmin(npp_prec, npp_temp)) %>% # kilograms of dry matter / square meter / year
  select(temperature, precipitation, npp) %>%
  st_crop(st_bbox(hex_geo)) %>%
  aggregate(hex_geo, mean, na.rm = TRUE) %>% 
  st_transform(hex) %>%
  st_as_sf()
```


```{r, message = FALSE}
elev <- read_stars('~/Downloads/EarthEnv-DEM90_N35E020/EarthEnv-DEM90_N35E020.bil') %>%
  st_crop(hex_geo)
```

```{r}
lake <- read_stars('~/kopais_105.tif') %>% 
  transmute(test = if_else(is.na( kopais_105.tif), NA_real_,1 )) %>% 
  st_as_sf(merge = TRUE)

elev2 <- (elev * ifelse(is.na(st_crop(elev, lake, crop = FALSE)[[1]]), 1, 0))

land_mask2 <- 
  aggregate(elev2 >= 1, hex_geo, mean, na.rm = TRUE) %>% 
  st_as_sf() %>%
  st_transform(st_crs(hex)) %>%
  rename(land_mask = EarthEnv.DEM90_N35E020.bil)

lake2 <- read_stars('~/kopais_95.tif') %>% 
  transmute(test = if_else(is.na( kopais_95.tif), NA_real_,1 )) %>% 
  st_as_sf(merge = TRUE)

elev3 <- (elev * ifelse(is.na(st_crop(elev, lake2, crop = FALSE)[[1]]), 1, 0))

land_mask3 <- aggregate(elev3 >= 1, hex_geo, mean, na.rm = TRUE) %>% 
  st_as_sf() %>%
  st_transform(st_crs(hex)) %>%
  rename(land_mask = EarthEnv.DEM90_N35E020.bil)

plot(land_mask);plot(land_mask2);plot(land_mask3)
```

```{rmessage = FALSE}
land_mask <- aggregate(elev >= 1, hex_geo, mean, na.rm = TRUE) %>% 
  st_as_sf() %>%
  st_transform(st_crs(hex)) %>%
  rename(land_mask = EarthEnv.DEM90_N35E020.bil)

arable <- elev %>%
  as('Raster') %>%
  raster::terrain(unit = 'degrees') %>%
  st_as_stars() %>%
  aggregate(hex_geo, FUN = function(x) sum(x <= 5, na.rm = TRUE) / length(x)) %>%
  st_transform(hex) %>% 
  st_as_sf() %>%
  st_join(land_mask, join = st_equals) %>%
  mutate(arable = as.numeric(slope * st_area(.) * land_mask),
         # mask out cells with less than half land (~8km2)
         arable = if_else(land_mask >= 0.5, arable, NA_real_))
```

multiply by to .25 to get aboveground biomass and grain. now edible grain in kg/m2, multiply times arable land to get total kg per cell, divide by 300 to get people it can support and * 0.50 for biennial fallow too. this is the free food from the agricultural population, so multiply by 0.1 as the "tax" that's actually available to move around
```{r, message = FALSE}
env <- st_join(arable, npp, join = st_equals) 

food <- transmute(env, carrying_capacity = npp * 0.25 * arable / 300 / 2 * 0.1)
```

```{r}
plot(env)
```

```{r}
plot(food)
```


```{r}
pts <- food %>%
  mutate(id = 1:n()) %>%
  filter(!is.na(carrying_capacity)) %>% 
  st_centroid()

plot(pts)
```

```{r}
settlements <- pts %>%
  st_distance() %>%
  units::set_units('km') %>%
  units::drop_units() %>% 
  replace(. == 0, 99999) %>%
  as_tbl_graph() %>%
  mutate(id = pts$id,
         population = 15,
         food = pts$carrying_capacity,
         eq = 0,
         x = st_coordinates(pts)[,1], 
         y = st_coordinates(pts)[,2]) %E>%
  rename(distance = weight) %>%
  # set self ties to half the cell radius, in km
  mutate(distance = if_else(distance == 99999, radius/2000, distance),
         trade_flow = 0,
         migrant_flow = 0)
```

```{r}
saveRDS(settlements, 'greece_settlements')
saveRDS(food, 'greece_env')
```


```{r}
settlements <- readRDS('greece_settlements')
food <- readRDS('greece_env')
```

get the lake free dem
```{r}
new_food <-  st_join(land_mask, food, join = st_equals) %>% mutate(new_car = carrying_capacity * land_mask) %>% select(new_car) %>% st_drop_geometry() %>% remove_missing()
```


select top ten cells, are how many of the city list are in there? -- that's the metric?

```{r}
target <- filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth'))
```

```{r}
calibrate <- function(x) {
  set.seed(x[1]) # necessary for multiocre ABC
  
  reduce(1:2000, ~run_sim(., beta1 = x[2], beta2 = x[3], alpha1 = x[4], alpha2 = x[5], nu = 0.05), 
                   .init = prune(settlements, max(x[2], x[3])))%>%
 as_tibble('nodes') %>%
  filter(population >= 1) %>%
 arrange(-population) %>%
  slice_head(n = 10) %>%
  st_as_sf(coords = c('x', 'y'), crs = 32634) %>%
  st_filter(target, .predicate = st_is_within_distance, dist = 5000) %>%
  nrow()
}

prior <- list(c('unif', 1, 20), c('unif', 1, 20), c('unif', 0.9, 1.2), c('unif', 0, 1.2))
prior.test = 'X1 <= X2'

abc_test <- EasyABC::ABC_rejection(calibrate, prior, 20, prior_test = 'X1 <= X2', summary_stat_target = 4, tol = 0.2, use_seed = TRUE, progress_bar = TRUE)

abc_test_seq <- EasyABC::ABC_sequential('Beaumont', calibrate, prior, 20, prior_test = 'X1 <= X2', summary_stat_target = 4, tolerance_tab = c(2, 1, .5), use_seed = TRUE, verbose = TRUE, progress_bar = TRUE)
```

```{r}
abc_test
```

```{r}
polis_sim <- accumulate(1:1000,
                   ~run_sim(., beta1 = 4, beta2 = 8, alpha1 = 1.1, alpha2 = 1, nu = 0.05), 
                   .init = prune(settlements, 8))
```

```{r}
read_csv('~/Downloads/points.csv')
```
```{r}
inv <- read_csv('~/Downloads/points.csv') %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  st_transform(32634) %>%
  st_crop(st_bbox(food))

ggplot(select(inv, Size)) + geom_sf(aes(size = Size, color = Size)) + geom_label_repel(data = filter(inv, Size >= 4), aes(geometry = geometry, label = name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps = 100) + scale_color_viridis_c() + scale_size_area()
```


```{r}
library(patchwork)

target <-poleis %>%
  filter(Name %in% c('Athens','Thebes', 'Argos', 'Korinth'))
site_names <- st_join(target, mutate(food, id = 1:n()), join = st_is_within_distance, dist = 8000) %>%
  as_tibble() %>%
  select(id, Name)



top10 <-   as_tibble(polis_sim[[1000]], 'nodes') %>%
    filter(population >= 1) %>%
  #  arrange(-population) %>% 
   #slice_head(n = 10) %>%
    st_as_sf(coords = c('x', 'y'), crs = 32634) %>%
    st_join(poleis, ., join = st_is_within_distance, dist = 6000, left = FALSE)

get_results(polis_sim[[1000]])
polis_sim[[1000]]
site_names()
```

```{r}
polis_sim %>%
  map_dfr(as_tibble, 'nodes', .id = 'time') %>%
  mutate(time = as.numeric(time)) %>%
  left_join(site_names) %>%
  ggplot(aes(time, population)) +
  geom_line(aes(group = id)) +
    gghighlight(Name %in% c('Athens','Thebes', 'Argos', 'Korinth'),
                label_key = Name,
                unhighlighted_params = list(alpha = 0.3), label_params = list(max.overlaps = 1000)) +
  theme_bw() +
  labs(x = 'Years', y = 'Population') 

polis_sim[[1000]] %>%
  filter(population >= 1) %>%
  ggraph(x = x, y = y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  #geom_edge_link(aes(color = migrant_flow)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area(name = 'Population') +
  geom_node_point(aes(size = population)) +
  theme_void() +
  
    geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), color = 'red') +
 # geom_label_repel(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)+
 # geom_sf(data = top10, color = 'orange') +
   # geom_label_repel(data = top10, aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps = 20) +
  geom_sf(data = select(inv, Size), aes(size = Size), color = 'white') + geom_label_repel(data = filter(inv, Size >= 3), aes(geometry = geometry, label = name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps = 100) + scale_color_viridis_c() +
  geom_sf(data = tibble(x=  690339 , y =4261700) %>% st_as_sf(coords = c('x','y'), crs =32634), color = 'purple', size = 2) +
  geom_sf(data = mycenaean %>% filter(Name != 'Thorikos: Hab'), color = 'red', aes(size = as.numeric(SizM)))
```
```{r}
mycenaean <- read_csv('analysis/mycenaean_atlas.csv') %>%
  mutate(Lat = as.numeric(Lat)) %>%
  remove_missing() %>%
  st_as_sf(coords = c('Lon', 'Lat'), crs = 4326) %>%
  st_transform(32634) %>%
  st_crop(st_bbox(food))# %>%
  ggplot() +
  geom_sf(aes(size = as.numeric(SizM)) )
  
  arrange(mycenaean, -as.numeric(SizM))
```


what do I want to do here . . . get the top ten sites on the dynamics?
```{r}
a <- polis_sim %>%
  map_dfr(as_tibble, 'nodes', .id = 'time') %>%
  mutate(time = as.numeric(time)) %>%
  left_join(site_names) %>%
  ggplot(aes(time, population)) +
  geom_line(aes(group = id)) +
    gghighlight(Name %in% c('Athens','Thebes', 'Argos', 'Korinth'),
                label_key = Name,
                unhighlighted_params = list(alpha = 0.3), label_params = list(max.overlaps = 1000)) +
  theme_bw() +
  labs(x = 'Years', y = 'Population')

b <- polis_sim[[932]] %>%
  filter(population >= 1) %>%
  ggraph(x = x, y = y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  #geom_edge_link(aes(color = migrant_flow)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area(name = 'Population') +
  geom_node_point(aes(size = population)) +
  theme_void() +
  
 #   geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth' , 'Kromna')), color = 'red') +
  geom_sf(data = target, color = 'orange') +
  geom_label_repel(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth', 'Merenda', 'Kromna')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)
```


```{r, fig.height=3.5, fig.width= 3.5}
b / a  + plot_layout(heights = c(2, 1))
```
```{r}
polis_sim[[350]] %>%
  filter(population >= 1) %>%
  ggraph(x = x, y = y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area(name = 'Population') +
  geom_node_point(aes(size = population)) +
    geom_sf(data = poleis, color = 'red') +

  theme_void() +
    geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth' )), color = 'red') +
  geom_sf(data = target, color = 'orange') +
  geom_label_repel(data = filter(poleis, Name %in% c('Potniai', 'Eutresis', 'Kabirion', 'Plataia', 'Thespiai', 'Thebes')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps =20)


polis_sim[[1000]] %>%
  filter(population >= 1) %>%
  ggraph(x = x, y = y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area(name = 'Population') +
  geom_node_point(aes(size = population)) +
  theme_void() +
    geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), color = 'red') +
  geom_sf(data = target, color = 'orange') +
    geom_label_repel(data = filter(poleis, Name %in% c('Potniai', 'Eutresis', 'Kabirion', 'Plataia', 'Thespiai', 'Thebes')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps =20)
```
```{r}

final_sites <-  as_tibble(polis_sim[[1000]], 'nodes') %>%
    filter(population >= 1) %>%
   # arrange(-population) %>% 
   #slice_head(n = 10) %>%
    st_as_sf(coords = c('x', 'y'), crs = 32634) 
polis_sim[[1000]] %>%
  filter(population >= 1) %>%
  ggraph(x = x, y = y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area(name = 'Population') +
  geom_node_point(aes(size = population)) +
  theme_void() +
   # geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), color = 'red') +
  geom_sf(data = st_filter(poleis, final_sites, .predicate = st_is_within_distance, dist = 4000), color = 'orange') +
    geom_label_repel(data = st_filter(poleis, final_sites, .predicate = st_is_within_distance, dist = 4000), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1, max.overlaps =20)

```


```{r}
st_filter(poleis, filter(poleis, Name == 'Thebes'), .predicate = st_is_within_distance, dist = 10000)
```



```{r}
#nd plots
polis_sim[[1000]] %>%
  nystuen_dacey() %>%
  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area() +
  geom_edge_link(aes(alpha = trade_flow)) +
  scale_edge_alpha() +
  geom_node_point(aes(size = population)) +
  theme_void()


polis_sim %>%
  nystuen_dacey(mode = 'migrants') %>%
  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area() +
  geom_edge_link(aes(alpha = migrant_flow)) +
  scale_edge_alpha() +
  geom_node_point(aes(size = population)) +
  theme_void()

```


```{r}

clust <-    polis_sim %E>%
  group_by(from) %>%
  filter(., trade_flow == max(trade_flow)) %N>%
  mutate(clust = group_components()) %>% pull(clust)


  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area() +
 # geom_edge_link() +
  geom_node_point(aes(size = population)) +
  theme_void() +
    geom_sf(data = poleis, color = 'red') +
  geom_label_repel(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)
  
  food %>% remove_missing() %>%
    mutate(clust = clust) %>%
  ggplot() +
  geom_sf(aes(fill = clust)) +
    scale_fill_viridis_c()

```

```{r}
polis_sim %E>% 
  group_by(from) %>%
  mutate(max_out = max(trade_flow)) %>%
  mutate(test = centrality_degree(weights = trade_flow, mode = 'in')>)

polis_sim %E>%
  group_by(from) %>%
  filter(trade_flow == max(trade_flow)) %>%
  ungroup %>%
  filter(trade_flow > .N()$harvest[from]) %N>%
  mutate(terminal = node_is_sink())
```


```{r}



polis_sim %>%
  filter(population >= 1) %>%
  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white', name = 'Local carrying capacity') +
  scale_size_area() +
  geom_node_point(aes(size = population)) +
  theme_void() +
    geom_sf(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth', 'Khalkis')), color = 'red') +
  #geom_sf(data = inv, color = 'orange') +
  geom_label_repel(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth', 'Khalkis')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)
```


```{r}
rm(polis_sim);gc()
polis_sim_cc <- accumulate(1:1000,
                   ~run_sim(., beta1 = 4, beta2 = 8, alpha1 = 1.1, alpha2 = 1, nu = 0.05),                    .init = prune(settlements, 8) %N>% mutate(population = c(rep(1, 510), 200, rep(1,181)))) 

                  # .init = prune(settlements, 8) %N>% mutate(population = food)) 
```

```{r}
polis_sim_cc %>%
  map_dfr(as_tibble, 'nodes', .id = 'time') %>%
  mutate(time = as.numeric(time)) %>%
  left_join(site_names) %>%
  ggplot(aes(time, population)) +
  geom_line(aes(group = id)) +
    gghighlight(Name %in% c('Athens','Thebes', 'Argos', 'Korinth', 'Khalkis'),
                label_key = Name,
                unhighlighted_params = list(alpha = 0.3)) +
  theme_bw() +
  labs(x = 'Years', y = 'Population')
```

```{r}
polis_sim_cc[[1000]] %>%
  filter(population >= 1) %>%
  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white') +
  scale_size_area() +
  geom_node_point(aes(size = population)) +
  theme_void() +
    geom_sf(data = poleis, color = 'red') +
  geom_label_repel(data = filter(poleis, Name %in% c('Athens','Thebes', 'Argos', 'Korinth')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)
```
```{r}
polis_sim_cc[[250]] %>%
  filter(population >= 1) %>%
  ggraph(x =x, y =y) +
  geom_sf(data = food, aes(fill = carrying_capacity)) +
  scale_fill_viridis_c(na.value = 'white') +
  scale_size_area() +
  geom_node_point(aes(size = population)) +
  theme_void() +
    geom_sf(data = poleis, color = 'red') +
  geom_label_repel(data = filter(poleis, Name %in% c('Orchomenos', 'Khalkis','Tiryns','Glisas','Kopai','Athens','Koropi','Tenea', 'Kromna','Thebes', 'Prof.Elias', 'Argos', 'Korinth')), aes(geometry = geometry, label = Name), stat = 'sf_coordinates',  box.padding = 1)
```

```{r}
polis_sim_cc[[1000]] %N>%
  as_tibble %>%
  left_join(site_names) %>%
  select(Name, population) %>%
  arrange(-population)
```


```{r}


polis_sim[[800]] %N>%
  as_tibble %>%
  filter(population > 1) %>%
  pull(population)  %>% hist
```



```{r}


test <- list.files('~/Downloads/rstudio-export-2', full.names = TRUE)[-1] %>%
  map_dfr(~readRDS(.) %>% mutate(test = map_dbl(sim, get_results)) %>%
  select(-sim))

test %>%
  filter(test >= 4) %>%
  select(-test) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_density() +
  #geom_histogram() +
  facet_wrap(~name, scales = 'free')
test %>%
  filter(test == 4) %>%
  select(-test) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_density() +
  #geom_histogram() +
  facet_wrap(~name, scales = 'free')
```

